Malacology: A Programmable Storage System

### 2016-10-20 paper/paper.tex
intro tweaks

A number of recent developments have disrupted traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
data'' applications demand interface evolution to support flexible consistency
as well as flexible structured data representations.
~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
storage systems available as open source software have established and are
continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

{\it \textbf{We define a programmable storage system}} to be a storage system
that facilitates the re-use and extension of existing storage abstractions
provided by the underlying software stack, to enable the creation of new
services via composition.  A programmable storage system can be realized by
exposing existing functionality (such as metadata services and synchronization
and monitoring capabilities) as interfaces that can be ``glued together'' in a
variety of ways using a high-level language. In contrast to programmable
storage, \emph{active storage} differs through its focus on the injection and
execution of code within a storage system or storage device. Programmable
storage differs from \emph{active storage}~\cite{riedel:vldb98} in that the
former is applicable to any component of the storage system, while the latter
focuses at the data access level (i.e. inject and execute arbitrary code at the
I/O access layers). Given this contrast, we can say that active storage is an
example of how one internal component (the storage layer) is exposed in a
programmable storage system.

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system that facilitates the
construction of new services by re-purposing existing subsystem abstractions of
the storage stack.  We build Malacology in Ceph, a popular open source software
storage stack.  We choose Ceph to demonstrate the concept of programmable
storage because it offers a broad spectrum of existing services, including
distributed locking and caching services provided by metadata servers,
durability and object interfaces provided by the backend object store, and
propagation of consistent cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

A prototype programmable storage system that includes a non-exhaustive set of
interfaces that can be used as building blocks for constructing novel storage
abstractions, including:

@@

\item A high-performance distributed shared log service called ZLog, that is an
implementation of CORFU~\cite{balakrishnan_corfu_2012}

@@

The remainder of this paper is structured as follows. First, we describe and
motivate the need for programmable storage by describing current practices in
the open source community. Next we describe Malacology by presenting the
subsystems within the underlying storage system that we re-purpose, and briefly
describe how those system are used within Malacology (\S\ref{sec:malacology}).
Then we describe the services that we have constructed within the Malacology
framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
implementation (\S\ref{sec:evaluation}).  Finally we discuss related work and
conclude.

>>>>

A number of recent developments have disrupted traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}.  Second, emerging ``big
data'' applications demand interface evolution to support flexible consistency
as well as flexible structured data 
representations.~\cite{apache_contributors_parquet_2014}.  Finally, production-quality scalable
storage systems available as open source software have established and are
continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

{\it \textbf{We define a programmable storage system}} to be a storage system
that facilitates the re-use and extension of existing storage abstractions
provided by the underlying software stack, to enable the creation of new
services via composition.  A programmable storage system can be realized by
exposing existing functionality (such as metadata services and synchronization
and monitoring capabilities) as interfaces that can be ``glued together'' in a
variety of ways using a high-level language. In contrast to programmable
storage, \emph{active storage} differs through its focus on the injection and
execution of code within a storage system or storage device. Programmable
storage differs from \emph{active storage}~\cite{riedel:vldb98}---the injection and
execution of code within a storage system or storage device---in
that the
former is applicable to any component of the storage system, while the latter
focuses at the data access level (i.e. inject and execute arbitrary code at the
I/O access layers). Given this contrast, we can say that active storage is an
example of how one internal component (the storage layer) is exposed in a
programmable storage system.

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system that facilitates the
construction of new services by re-purposing existing subsystem abstractions of
the storage stack.  We build Malacology in Ceph, a popular open source software
storage stack.  We choose Ceph to demonstrate the concept of programmable
storage because it offers a broad spectrum of existing services, including
distributed locking and caching services provided by metadata servers,
durability and object interfaces provided by the backend object store, and
propagation of consistent cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
expressive enough to provide the functionality necessary for implementing new
services. 
%Our contributions are:

% prototype programmable storage system that 
Malacology
includes a non-exhaustive set of
interfaces that can be used as building blocks for constructing novel storage
abstractions, including:

@@

\item A high-performance distributed shared log service called ZLog, that is an
implementation of CORFU.~\cite{balakrishnan_corfu_2012}

@@

The remainder of this paper is structured as follows. First, we describe and
motivate the need for programmable storage by describing current practices in
the open source community. Next we describe Malacology by presenting the
subsystems within the underlying storage system that we re-purpose, and briefly
describe how those system are used within Malacology (\S\ref{sec:malacology}).
Then we describe the services that we have constructed within the Malacology
framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
implementation (\S\ref{sec:evaluation}).  We conclude by discussing related and future work.
% and
%conclude.
